The \frb Observing Campaign: Multi-Telescope Radio Burst and Implications for the FRB Population

### 2016-12-17 repeater-vladata.tex
added dm confidence interval figure. working on non-lambda2 fits

\title{\frb: Multi-Telescope Burst Properties and Implications for FRB Population}
\shorttitle{\frb\ Burst Properties}
\shortauthors{Law et al.}

@@

This paper is part of a series that presents the first localization and unambiguous identification of an FRB host (CITE: LOC, OPT, EVN, RADIO CPART, XRAY, EFF, WEIRD). \frb, also known as the ``repeating FRB'', was first detected in November 2012 by the Arecibo Observatory \citep{2014ApJ...790..101S}. In mid 2015, new Arecibo observations revealed a series of bursts at the same DM and sky position demonstrating that FRBs are capable of repetition \citep{2016Natur.531..202S}. Beginning in August of 2015, we made the first of nine detections of \frb\ with the Very Large Array (CITE: LOC) and localized it with a precision of 0.1\arcsec. Deep radio and optical observing shows that \frb\ is unambiguously associated with a persistent radio and optical source at a redshift of 0.193 (CITE: OPT, EVN).

\frb\ has now been localized three orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs are more luminous than any other millisecond radio transient, have a significant DM contribution from the IGM, and can be used to probe the IGM and their host galaxy. The promise implied by the first reported FRB \citep{2007Sci...318..777L} is now being realized.

@@

The data presented here were obtained from multiple programs and telescopes, but the central goal was to interferometrically localize \frb\ with the VLA. The observing strategy was to ensure simultaneous observing between the VLA and Arecibo observatories and add other observatories added on a best-effort basis. We coordinated observing between the VLA, Arecibo, Effelsberg, and AMI-LA telescopes, as shown in figure \ref{fig:multi}. Below, we summarize these observations, with a focus on those conducted simultaneous with VLA burst detections from \frb.

\subsection{VLA}
The \frb\ observing campaign started in late 2015 with a 10~hr campaign observed at 1.4~GHz in the compact D configuration. In April through May 2016, we conducted a 40~hr campaign at 3~GHz in the C and CnB configurations in coordination with Arecibo. We concluded with a new, 40~hr, coordinated campaign from August through September 2016 in the B configuration and during the move to the most extended A configuration. In this last campaign, the first 34 hours of VLA observations were made at 3~GHz, while the last 6 hours were observed at 6~GHz. This paper focuses on the data collected at 3~GHz, which includes all burst detections.

All VLA fast-sampled data were observed with 5~ms sampling, 256 channels, and dual-circular polarization \citep[as in]{2015ApJ...807...16L}. To maximize sensitivity, the channel frequency width was set to maintain sensitivity to the known DM of \frb, while maximizing the total bandwidth. The total bandwidth at L (1.4~GHz), S (3~GHz), and C (6~GHz) bands was 256~MHz, 1024~MHz, and 2048~MHz, respectively. The 3~GHz data recorded data in 8 spectral windows with 32 channels each.

Observations in August and Septmeber were searched by a prototype version of \rf\footnote{See \url{http://realfast.io}.}. \rf\ is a real-time, fast imaging transient search system. The current, prototype runs on existing, CPU-based hardware of the VLA correlator backend, while the future \rf\ will run on a dedicated GPU cluster. The transient search pipeline software is called \emph{rtpipe}\footnote{See \url{https://github.com/caseyjlaw/rtpipe}} and is mostly written in Python. The transient search pipeline is described in \citet{2015ApJ...807...16L} and performs data calibration, flagging, dedispersion, and imaging. Images were formed for each integration at DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$. Gain calibration is read from the "telcal" system, which uses phase-only calibration on the previous gain calibrator. A flux scale is calculated for each spectral window from an observation on {\color{red} xx (Bryan)} and applied to all burst spectra.

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf\ is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". Each image with a pixel higher than 6.4$\sigma$ has some rudimentary data saved to capture both real and thermal-noise candidates as a check of data quality. For each image with a pixel higher than 7.4$\sigma$, \rf\ generates a candidate visualization with an image and spectrum. More detailed analysis, including improved calibration and localization, is conducted offline.

@@

(copied from the first paper) 
During the joint Arecibo-VLA campaign, Arecibo observed with the L-wide receiver, which has an observational frequency range of 1.15 to 1.73~GHz and a full width at half maximum beam size of 3.3 arcmin. The PUPPI pulsar backend was used to record total intensity spectra with time and frequency resolutions of 10.24~$\mu$s and 1.5625~MHz, respectively, and full Stokes polarization information. Each frequency channel was coherently dedispersed to 557 pc cm$^{-3}$, thereby eliminating intra-channel dispersion smearing. PUPPI covers a total of 800~MHz of bandwidth centred at 1380.78125~MHz, but only $\sim$ 620~MHz of this band is usable due to radio frequency interference and receiver sensitivity roll-off at the band edges.

In total, twelve Arecibo observations had some simultaneous coverage with the VLA. Four of those observations were simultaneous with bursts detected with the VLA and one of those observations detected the same VLA burst. During the first VLA burst with Arecibo coverage (MJD 57643), the PUPPI system failed so data were recorded with {\color{red} xx (Jason?)} at C band. No detection was made in that Arecibo data. Overall, there were many more bursts detected at Arecibo and a more detailed analysis will be presented in a future paper.

@@

(copied from Laura's paper)
Observations were conducted at an observing frequency of 4.6 to 5.1~GHz. The S60mm receiver has a system equivalent flux density of 18 Jy and a full-width half-max (FWHM) beam size of 2.4\arcmin at 4.85~GHz. Pulsar search mode data were recorded with the PFFTS backend. Total intensity spectra were recorded with a time resolution of 65.5~$\mu$s and a bandwidth of 500~MHz divided into 128 frequency channels. Note, the inter-channel DM smearing time for 560~pc cm$^{-3}$\ is $\sim$0.2 msec at 4.6 GHz.

@@

We observed FRB121102 with the Arcminute MicroKelvin Imager Large Array (AMI-LA; Zwart et al. 2008) for 3 hours each on four epochs starting MJDs 57643.3351, 57645.3317, 57648.3352, 57649.3353. Observations were made with the new digital correlator having 4096 channels across a 5 GHz bandwidth between 13--18 GHz with a 1s dump time. The phase calibrator, J0518+3306, was observed every 12 minutes for about 1.5 minutes. The AMI-LA data were binned to eight 0.625 GHz channels and processed (RFI excision and calibration) with a fully-automated pipeline, AMI-REDUCE (e.g. Davies et al. 2009; Perrott et al. 2013). Daily measurements of 3C48 and 3C286 were used for the absolute flux calibration, which is good to about 10\%.

**find a place for this elsewhere**
We inspected the calibrated visibilities, and did not find any signal above 20 mJy in the 1s samples at and in the vicinity of the detected bursts (MJDs 57643.45730263, 57645.42958602, 57648.4369149, and 57649.45175697 **add two more with VLA coverage**). Concatenating and imaging the 12 hours of calibrated data with the CASA tasks {\it concat} and {\it clean} also does not yield any significant detection at the FRB location. Although the statistical 3sigma upper limit is 60 uJy, extended mJy-level sources in the field cause sidelobe confusion (the AMI-LA angular resolution is $\sim$30 arcsec), and the actual upper limit is larger. We introduced artificial point sources at the FRB location using the CASA {\it sm} tool, and found that these sources can be recovered as long as their peak flux densities are more than $\sim$100 uJy. Hence, we place an upper limit of 100$\pm$10 uJy on any quiescent or possible radio "afterglow" (on $\sim$days timescale) signal from the FRB.

Deep image sensitivity is limited by confusion and is approximately equal to the flux density observed by VLA (CITE: RADIO CPAR

@@

\subsection{Multi-Telescope Burst Spectrum}
Of the four VLA burst detections from \frb with simultaneous observations \frb (Figure \ref{fig:multi}), one was detected with a consistent dedispersed time. The VLA burst on 57648 was detected at Arecibo with a significance of {\color{red} xx$\sigma$ (Jason)} The relative time delay between the VLA and Arecibo is {\color{red} xx (Jason)} ms...

Figure **to do** shows the dynamic spectrum formed from the phased VLA and Arecibo data... 
This simultaneous detection of a burst from roughly 1 to 4~GHz shows that some bursts cover more than an octave of frequency. This is one of three bursts with similar spectral coverage at Arecibo, so the other nondetections imply that those bursts had a smaller spectral coverage. In fact, as described in \S \ref{sec:spec}, most of the VLA bursts appear to be fully contained in the band from 2.5 to 3.5~GHz, which implies a much smaller spectral coverage for typical bursts.

**Insert figure of VLA/AO dynamic spectrum on 57648**

@@

 \caption{Spectrograms (time vs frequency intensity maps) for the nine VLA bursts. Starting at the top left, the correspond to bursts 57623, 57633.68, 57633.70, 57638, 57643, 57645, 57646, 57648, and 57649. Note that bursts are detected in 5~ms images generated from dedispersed visibilities.
 \label{fig:sgram}}
\end{center}
\end{figure*}

\subsubsection{Dynamic Spectra}
\label{sec:spec}
Figure \ref{fig:sgram} shows the spectrograms of all nine bursts detected by the VLA. We improved on the initial analysis presented in (CITE: LOC) by using a better calibration scheme and optimizing the detection signifnance by optimizing over a range of DM. Table \ref{tab:spec} shows the improved burst parameters using this new scheme.

@@

\subsubsection{Dispersion Meausres}
The initial VLA detections were made by searching a DM grid that allowed inter-DM sensitivity losses up to roughly 10\%. For the VLA 3~GHz observing band, we estimated that trials could be separated by at most 10 pc cm$^{-3}$ \citep{2003ApJ...596.1142C}. We refined that analysis by maximizing the image SNR for DMs spaced by 1 pc cm$^{-3}$. The optimal DM values range from 552 to 572 pc cm$^{-3}$ for all bursts and from 552 to 561 pc cm$^{-3}$ for bursts with SNR$>30$.

The uncertainty in the peak DM measurement is defined by the DM sensitivity loss curve and the significance of the burst. We tested the precision of the peak DM measuremet by simulating a burst with SNR$=30$ and measuring its detection significance as a function of DM. The simulation parameters were reproduced at the known DM with a precision of roughly 1 pc cm$^{-3}$. This suggests that the measured variation in DM for \frb is intrinsic to the bursts, rather than due to measurement uncertainty. This is consistent with the observation of intrinsic burst spectrotemporal structure that varies between bursts (CITE: WEIRD).

\subsubsection{Spectral modeling}
After finding an optimal burst DM, we extract a spectrum from the integration with peak SNR (Figure \ref{fig:spec}). After DM optimization, all bursts appear unresolved at the 5~ms time resolution of the VLA data (except maybe 57633.68, brightest burst).

>>>>

\title{The \frb Observing Campaign: Multi-Telescope Radio Burst and Implications for the FRB Population}
%\title{Multi-Telescope Burst Properties and Implications for FRB Population}
\shorttitle{\frb\ Observing Campaign}
\shortauthors{Law et al.}

@@

This paper is part of a series that presents the first localization and unambiguous identification of an FRB host (CITE: LOC, OPT, EVN). \frb, also known as the ``repeating FRB'', was first detected in November 2012 by the Arecibo Observatory \citep{2014ApJ...790..101S}. In mid 2015, new Arecibo observations revealed a series of bursts at the same DM and sky position demonstrating that FRBs are capable of repetition \citep{2016Natur.531..202S}. Beginning in August of 2015, we made the first of nine detections of \frb\ with the Very Large Array (CITE: LOC) and localized it with a precision of 0.1\arcsec. Deep radio and optical observing shows that \frb\ is unambiguously associated with a persistent radio and optical source at a redshift of 0.193 (CITE: OPT, EVN).

\frb\ has now been localized four orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs are more luminous than any other millisecond radio transient, have a significant DM contribution from the IGM, and can be used to probe the IGM and their host galaxy. The promise implied by the first reported FRB \citep{2007Sci...318..777L} is now being realized.

@@

The data presented here were obtained from multiple programs and telescopes, but the central goal was to interferometrically localize \frb\ with the VLA. The observing strategy was to ensure simultaneous observing between the VLA and Arecibo observatories and add other observatories added on a best-effort basis. We coordinated observing between the VLA, Arecibo, Effelsberg, and AMI-LA telescopes, as shown in figure \ref{fig:multi}. Below, we summarize these observations, with a focus on those conducted simultaneously with VLA burst detections from \frb.

\subsection{VLA}
The \frb\ observing campaign started in late 2015 with a 10~hr campaign observed at 1.4~GHz in the compact D configuration. In April through May 2016, we conducted a 40~hr campaign at 3~GHz in the C and CnB configurations in coordination with Arecibo. We concluded with a new, 40~hr, coordinated campaign from August through September 2016 in the B configuration and during the move to the most extended A configuration. In this last campaign, the first 34 hours of VLA observations were made at 3~GHz, while the last 6 hours were observed at 6~GHz. This paper focuses on the data collected at 3~GHz, which includes all nine burst detections.

All VLA fast-sampled data were observed with 5~ms sampling, 256 channels, and dual-circular polarization \citep[as in]{2015ApJ...807...16L}. To maximize sensitivity, the channel frequency width was set to maintain sensitivity to the known DM of \frb, while maximizing the total bandwidth. The total bandwidth at L (1.4~GHz), S (3~GHz), and C (6~GHz) bands was 256~MHz, 1024~MHz, and 2048~MHz, respectively. The 3~GHz data were recorded data in 8 spectral windows with 32 channels each.

Observations in August and Septmeber were searched by a prototype version of \rf\footnote{See \url{http://realfast.io}.}. \rf\ is a real-time, fast imaging transient search system. The current, prototype runs on existing, CPU-based hardware of the VLA correlator backend, while the future \rf\ will run on a dedicated GPU cluster. The transient search pipeline software is called \emph{rtpipe}\footnote{See \url{https://github.com/caseyjlaw/rtpipe}} and is mostly written in Python (detail in \citet{2015ApJ...807...16L}). Images were formed for each integration at DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$. Gain calibration is read from the "telcal" system, which uses phase-only calibration on the previous gain calibrator. A flux scale is calculated for each spectral window from an observation on {\color{red} xx (Bryan)} and applied to all burst spectra.

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf\ is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". Each image with a pixel higher than 6.4$\sigma$ is saved in a summary form as a check of data quality. For each image with a pixel higher than 7.4$\sigma$, \rf\ generates a more detailed candidate visualization with an image and spectrum. More detailed analysis, including improved calibration and localization, is conducted offline.

@@

{\color{red} copied from the first paper}
During the joint Arecibo-VLA campaign, Arecibo observed with the L-wide receiver, which has an observational frequency range of 1.15 to 1.73~GHz and a full width at half maximum beam size of 3.3 arcmin. The PUPPI pulsar backend was used to record total intensity spectra with time and frequency resolutions of 10.24~$\mu$s and 1.5625~MHz, respectively, and full Stokes polarization information. Each frequency channel was coherently dedispersed to 557 pc cm$^{-3}$, thereby eliminating intra-channel dispersion smearing. PUPPI covers a total of 800~MHz of bandwidth centred at 1380.78125~MHz, but only $\sim$ 620~MHz of this band is usable due to radio frequency interference and receiver sensitivity roll-off at the band edges.

In total, twelve Arecibo observations had some simultaneous coverage with the VLA. Four of those observations were simultaneous with bursts detected with the VLA and one of those observations detected the same VLA burst. During the first VLA burst with Arecibo coverage (MJD 57643), the PUPPI system failed so data were recorded with {\color{red} xx (Jason?)} at C band. No detection was made in that Arecibo data. Overall, there were many more bursts detected at Arecibo than with the VLA and a more detailed analysis of those bursts will be presented in a future paper.

@@

{\color{red} copied from Laura's paper}
Observations were conducted at an observing frequency of 4.6 to 5.1~GHz. The S60mm receiver has a system equivalent flux density of 18 Jy and a full-width half-max (FWHM) beam size of 2.4\arcmin at 4.85~GHz. Pulsar search mode data were recorded with the PFFTS backend. Total intensity spectra were recorded with a time resolution of 65.5~$\mu$s and a bandwidth of 500~MHz divided into 128 frequency channels. Note, the inter-channel DM smearing time for 560~pc cm$^{-3}$\ is $\sim$0.2 msec at 4.6 GHz.

@@

We observed \frb with the Arcminute MicroKelvin Imager Large Array (AMI-LA; Zwart et al. 2008) for 3 hours each on four epochs starting at MJDs 57643.3351, 57645.3317, 57648.3352, and 57649.3353. Observations were made with the new digital correlator having 4096 channels across a 5 GHz bandwidth between 13--18 GHz with a 1s dump time. The phase calibrator, J0518+3306, was observed every 12 minutes for about 1.5 minutes. The AMI-LA data were binned to eight 0.625 GHz channels and processed (RFI excision and calibration) with a fully-automated pipeline, AMI-REDUCE \citep[e.g.,][]{2013MNRAS.429.3330P}. Daily measurements of 3C48 and 3C286 were used for the absolute flux calibration, which is good to about 10\%.

We inspected the calibrated visibilities, and did not find any signal above 20 mJy in the 1s samples at and in the vicinity of the detected bursts. Concatenating and imaging the 12 hours of calibrated data with the CASA tasks {\it concat} and {\it clean} also does not yield any significant detection at the FRB location. Although the statistical 3sigma upper limit is 60 uJy, extended mJy-level sources in the field cause sidelobe confusion (the AMI-LA angular resolution is $\sim$30 arcsec), and the actual upper limit is larger. We introduced artificial point sources at the FRB location using the CASA {\it sm} tool, and found that these sources can be recovered as long as their peak flux densities are more than $\sim$100 uJy. Hence, we place an upper limit of 100$\pm$10 uJy on any quiescent or possible radio "afterglow" (on $\sim$days timescale) signal from the FRB. This limit is close to the flux density measured by the VLA (CITE: LOC)

@@

\subsection{Multi-Telescope Burst Spectrum}
Four of the VLA bursts were observed simultaneously with Arecibo and AMI-LA (Figure \ref{fig:multi}). Of these, the burst on 57648 was detected by Arecibo at the same time (after correcting for the known DM of \frb). The Arecibo detection of this burst had a significance of {\color{red} xx$\sigma$ (Jason)}.

Figure {\color{red} **to do**} shows the dynamic spectrum formed from the phased VLA and Arecibo data... **Discuss dispersion correction, time alignment, chance of false association**

This simultaneous detection of a burst from roughly 1 to 4~GHz shows that some bursts cover more than an octave of frequency. However, this is one of three VLA bursts with coverage by Arecibo, so the other nondetections imply that those bursts had a smaller spectral width. In fact, as described in \S \ref{sec:spec}, many of the VLA bursts appear to be fully contained in the band from 2.5 to 3.5~GHz, which implies a much smaller spectral coverage for typical bursts.

@@

 \caption{Dynamic spectra (time vs frequency intensity maps) for the nine VLA bursts. Starting at the top left, they correspond to bursts 57623, 57633.68, 57633.70, 57638, 57643, 57645, 57646, 57648, and 57649. Note that bursts are detected in 5~ms images generated from dedispersed visibilities.
 \label{fig:sgram}}
\end{center}
\end{figure*}

\subsubsection{Dynamic Spectra}
\label{sec:spec}
Figure \ref{fig:sgram} shows the dynamic spectra of all nine bursts detected by the VLA. We improved on the initial analysis presented in (CITE: LOC) by using a better calibration scheme and optimizing the detection signifnance over a range of DM. Table \ref{tab:spec} shows the improved burst parameters using this new scheme.

@@

\subsubsection{Dispersion Meausres}
The initial VLA detections were made by searching a DM grid that allowed inter-DM sensitivity losses up to roughly 10\%. For the VLA 3~GHz observing band, we estimated that trials could be separated by at most 10 pc cm$^{-3}$ \citep{2003ApJ...596.1142C}. We refined that analysis by maximizing the image SNR for DMs spaced by 1 pc cm$^{-3}$. The optimal DM values range from 552 to 572 pc cm$^{-3}$.

The uncertainty in the peak DM measurement is defined by the DM sensitivity loss curve and the significance of the burst. We estimate the DM likelihood distribution for each burst by using a generative model for the burst spectra \citep{2010arXiv1008.4686H}. The burst spectra as a function of frequency were fixed according to the broad Gaussian shape defined below (\S \ref{sec:specmod}) and made into dynamic spectra by applying a frequency-dependent delay. We assume a Gaussian error distribution for channels with low signal and error up to 100\% of the model amplitude for channels with high signal. This is consistent with the large channel-to-channel variations observed (potentially due to diffractive scintillation).

Figure \ref{fig:dmmodel} compares the 68\% ($1\sigma$) bounds on the DM for all nine VLA bursts.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{dmmodel}
\caption{Comparison of 68\% ($\1sigma$) confidence intervals for DM for nine bursts detected by the VLA.
\label{fig:dmmodel}}
\end{center}
\end{figure}

\subsubsection{Spectral modeling}
\label{sec:specmod}
After finding an optimal burst DM, we extract a spectrum from the integration with peak SNR (Figure \ref{fig:spec}). After DM optimization, all bursts appear unresolved at the 5~ms time resolution of the VLA data (except maybe 57633.68, brightest burst).
